---
title: "`greet` app for 'Getting Started with {`shinytest2`}'"
output: 
  github_document:
    toc: true
    toc_depth: 4

---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE, 
  comment = "#>>", 
  collapse = TRUE)
```

# `shinytest2` [part 1](https://www.youtube.com/watch?v=Gucwz865aqQ)

**Example + basics**

<!-- [HackMD](https://hackmd.io/@mfrigaard/Syag4ij72) -->

1. Install Chromium via [homebrew](https://formulae.brew.sh/cask/chromium) and dev version of `chromote`

```{bash}
#| eval: false
==> Downloading https://download-chromium.appspot.com/dl/Mac?type=snapshots
==> Downloading from https://commondatastorage.googleapis.com/chromium-browser-s
######################################################################### 100.0%
Warning: No checksum defined for cask 'chromium', skipping verification.
==> Installing Cask chromium
==> Moving App 'Chromium.app' to '/Applications/Chromium.app'
==> Linking Binary 'chromium.wrapper.sh' to '/usr/local/bin/chromium'
🍺  chromium was successfully installed!
``` 

```{r}
#| eval: false
remotes::install_github("rstudio/chromote")
library(chromote)
```

2. Located Chromium 

```{r}
#| eval: false
chromote::find_chrome()
# [1] "/Applications/Chromium.app/Contents/MacOS/Chromium"
```

3. Created new session 

```{r}
#| eval: false
Sys.setenv(CHROMOTE_CHROME = "/Applications/Chromium.app/Contents/MacOS/Chromium")
b <- ChromoteSession$new()
b$view()
```


## The application 

From [this example](https://www.youtube.com/watch?v=Gucwz865aqQ)

```{r}
#| eval: false
library(shiny)
library(magrittr)

ui <- fluidPage(
  textInput("name", "what is your name"),
  actionButton("greet", "Greet"),
  textOutput("greeting"),
  textOutput("first_letter")
)

server <- function(input, output, session) {
    
  output$greeting <- renderText({
    req(input$name)
    paste0("Hello ", input$name, "!")
  }) %>% 
    bindEvent({input$greet})
    
  first_letter <- reactive({
    req(input$name)
    tolower(stringr::str_extract(input$name, "^."))
  }) %>% 
    bindEvent({input$greet})
  
  output$first_letter <- renderText({
    paste0("The first letter in your name is ", first_letter(), "!")
  })
  
}

shinyApp(ui, server)
```

## Record test

```{r}
#| eval: false
shinytest2::record_test()
```

1. Enter a name and click "Greet"

2. Enter name for test (i.e., `hello-barret`) 

3. Click **Expect Shiny values**

4. Save and exit

## Test

1. The following folders/files are created:

```
├── app.R
├── greet.Rproj
└── tests
    ├── testthat
    │   ├── _snaps
    │   │   └── shinytest2
    │   │       ├── hello-barret-001.json
    │   │       └── hello-barret-001_.png
    │   ├── setup-shinytest2.R
    │   └── test-shinytest2.R
    └── testthat.R

5 directories, 7 files
```

2. Contents of `test-shinytest2.R`:

```{r}
#| eval: false
library(shinytest2)

test_that("{shinytest2} recording: hello-barret", {
  app <- AppDriver$new(name = "hello-barret", height = 483, width = 862)
  app$set_inputs(name = "Barret")
  app$click("greet")
  app$expect_values()
})
```

## Debugging in console

Create `app` with `AppDriver$new()` in the console

```{r}
#| eval: false
app <- AppDriver$new(name = "hello-barret", height = 483, width = 862)
```

```{r}
#| eval: false
app$view()
```

*Opens Chromium with app*

```{r}
#| eval: false
app$set_inputs(name = "Barret")
```

*Sets value in `input$name`*

```{r}
#| eval: false
app$click("greet")
```

*Runs `actionButton(input$click)`*

```{r}
#| eval: false
app$get_values()
```

```
#>> $input
#>> $input$greet
#>> [1] 2
#>> attr(,"class")
#>> [1] "shinyActionButtonValue" "integer"               

#>> $input$name
#>> [1] "Barret"


#>> $output
#>> $output$first_letter
#>> [1] "The first letter in your is b!"

#>> $output$greeting
#>> [1] "Hello Barret!"


#>> $export
#>> named list()
```

```{r}
#| eval: false
app$get_value(output = "greeting")
```

```
#>> [1] "Hello Barret!"
```


# `shinytest2` [part 2](https://www.youtube.com/watch?v=7KLv6HdIxvU)

**Exporting values**

1. Create new test in `test-shinytest2.R`

-   This doesn't require `height = 483, width = 862`, but it makes the size eaiser to deal with. 

    -   Add the `first_letter_phrase` as the output from `app$get_value(output = "first_letter")`
    
    -   Create test using the `first_letter_phrase` and the expected output in `output$first_letter`

    ```{r}
    #| eval: false
    testthat::test_that(desc = "export test values", {
      app <- AppDriver$new(name = "export values")
      app$set_inputs(name = "Barret")
      app$click("greet")
      first_letter_phrase <- app$get_value(output = "first_letter")
      testthat::expect_equal(
        object = first_letter_phrase,
        expected = "The first letter in your name is b!")
    })
    ```
    
### Exporting values 

To export values, we can use the `shiny::exportTestValues()` in the `server`: 

```{r}
#| eval: false
shiny::exportTestValues(
  first_letter =
    first_letter()
  )
```


-   Then access this in the `AppDriver`: 

    -   Reset the values 

    ```{r}
    #| eval: false
    app$stop()
    app <- AppDriver$new(name = "export values")
    app$view()
    app$set_inputs(name = "Barret")
    app$click("greet")
    ```
    
    -   View exported values 
    
    ```{r}
    #| eval: false
    app$get_values()
    ```
    
    -   Test exported values 
    
    ```{r}
    test_that(desc = "export test values", code = {
      app <- AppDriver$new(name = "export values")
      app$set_inputs(name = "Barret")
      app$click("greet")
    
      # extract value from output$first_letter
      first_letter_phrase <- app$get_value(output = "first_letter")
    
      expect_equal(object = first_letter_phrase,
        expected = "The first letter in your name is b!")
    
      expect_equal(
        object = app$get_value(export = "first_letter"),
        expected = "b"
      )
    
    })
    ```
    

# `shinytest2` [part 3](https://www.youtube.com/watch?v=xDxa_mDwN04)

**Using `shiny.testmode`**

1. Update the `ui` with a `textOutput()`

```{r}
#| eval: false
textOutput("penguins")
```

2. Update the server with a simulated `authenticated_database_request()`

```{r}
#| eval: false
authenticated_database_request <- isolate({
    reactive({
      data.frame(x = vector(
        mode = "integer",
        length = sample(
          x = c(320:340),
          size = 1,
          replace = TRUE
        )
      ))
    })
  }) %>%
    bindEvent({input$greet})
```

-   This is used in `penguin_count()`

    ```{r}
    #| eval: false
     penguin_count <- reactive({
        penguin_data <- authenticated_database_request()
        nrow(penguin_data)})  %>%
        bindEvent({input$greet})
    ```
    
    -   The output is below: 
    
    ```{r}
    #| eval: false
    output$penguins <- renderText({
      paste0("There are ", penguin_count(), " penguins coming for dinner")
    })
    ```
    
When we use `shiny.testmode`, we add the simulated `test_database_request()` to the control flow for `penguins_count()`: 

-   The `test_database_request()`

    ```{r}
    #| eval: false
     test_database_request <- isolate({
        reactive({
          data.frame(x = vector(
            mode = "integer",
            length = 68L
          ))
        })
      }) %>%
        bindEvent({input$greet})
    ```
    
    -   The updated server with `penguins_count()`
    
    ```{r}
    #| eval: false
    penguin_count <- reactive({
      penguin_data <-
      if (getOption("shiny.testmode", FALSE)) {
        test_database_request()
      } else {
        authenticated_database_request()
      }
      nrow(penguin_data)
    })  %>%
      bindEvent({input$greet})
    ```
    
    -   These values can also be exported: 
    
    ```{r}
    #| eval: false
    shiny::exportTestValues(
        first_letter =
          first_letter(),
        penguin_count =
          penguin_count()
      )
    ```
    
### Test with `shiny.testmode`

After resetting the `AppDriver`, we can test for the `test_database_request()` using the following in the console:

```{r}
#| eval: false
app <- AppDriver$new(name = "export values", height = 483, width = 862) 
app$set_inputs(name = "Barret")
app$click("greet")
app$get_value(export = "penguin_count")
#>> [1] 68
```

-   And add the following to `test-shinytest2.R`: 

    ```{r}
    #| eval: false
      testthat::expect_equal(
        object = app$get_value(export = "penguin_count"),
        expected = 68
      )
    ```

# Final `test-shinytest2.R`

```{r}
#| eval: false
library(testthat)
library(shinytest2)

# initial record_test() ----
# testthat::test_that("{shinytest2} recording: hello-barret", {
#   app <- AppDriver$new(name = "hello-barret", height = 483, width = 862)
#   app$set_inputs(name = "Barret")
#   app$click("greet")
#   app$expect_values()
# })

testthat::test_that(desc = "export test values", code = {
  
  app <- AppDriver$new(name = "export values", height = 483, width = 862) 
  app$set_inputs(name = "Barret")
  app$click("greet")
  
  first_letter_phrase <- app$get_value(output = "first_letter")
  
  testthat::expect_equal(object = first_letter_phrase, 
    expected = "The first letter in your name is b!")
  
  testthat::expect_equal(
    object = app$get_value(export = "first_letter"),
    expected = "b"
  )
  
  testthat::expect_equal(
    object = app$get_value(export = "penguin_count"),
    expected = 68
  )
  
})


```


# Final `app.R`

```{r}
library(shiny)
library(magrittr)

ui <- fluidPage(
  textInput("name", "what is your name"),
  actionButton("greet", "Greet"),
  textOutput("greeting"),
  textOutput("first_letter"),
  # penguins ----
  textOutput("penguins")
)

server <- function(input, output, session) {
  
  # authenticated_database_request ----
  authenticated_database_request <- isolate({
    reactive({
      data.frame(x = vector(
        mode = "integer",
        length = sample(
          x = c(320:340),
          size = 1,
          replace = TRUE
        )
      ))
    })
  }) %>%
    bindEvent({
      input$greet
    })

  # test_database_request ----
  test_database_request <- isolate({
    reactive({
      data.frame(x = vector(
        mode = "integer",
        length = 68L
      ))
    })
  }) %>%
    bindEvent({
      input$greet
    })


  output$greeting <- renderText({
    req(input$name)
    paste0("Hello ", input$name, "!")
  }) %>%
    bindEvent({
      input$greet
    })

  first_letter <- reactive({
    req(input$name)
    tolower(stringr::str_extract(input$name, "^."))
  }) %>%
    bindEvent({
      input$greet
    })

  # penguin_count ----
  penguin_count <- reactive({
    penguin_data <-
      if (getOption("shiny.testmode", FALSE)) {
        test_database_request()
      } else {
        authenticated_database_request()
      }
    nrow(penguin_data)
  }) %>%
    bindEvent({
      input$greet
    })

  output$first_letter <- renderText({
    paste0("The first letter in your name is ", first_letter(), "!")
  })

  output$penguins <- renderText({
    paste0("There are ", penguin_count(), " penguins coming for dinner")
  })

  # exportTestValues ----
  shiny::exportTestValues(
    first_letter =
      first_letter(),
    penguin_count =
      penguin_count()
  )
}

shinyApp(ui, server)

```

